---
title: "Reproducible Research: Peer Assessment 1"
html_document:
keep_md: true
output: github_document
---

<div style="margin-bottom:20px;">
</div>

***27/03/2020***

<div style="margin-bottom:40px;">
</div>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.path='Figs/')
library(dplyr)
```

## Data handling

### Some code for downloading and reading the data

```{r reading data}
url1<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
down<-download.file(url = url1,"data.zip")
file1<-unzip("data.zip")
ActivityData<-read.csv(file =file1)
```

### Some code for tranforming the data

```{r transformations}
ActivityData$date<-as.Date(ActivityData$date)
```

<div style="margin-bottom:30px;">
</div>


## **Steps per day**

### Total number of steps per day

```{r totalSteps per day,fig.width=3}
StepsPerDay<-tapply(ActivityData$steps,ActivityData$date,sum)
data.frame(StepsPerDay)
```

### Histogram of total steps taken per day

```{r hist totalSteps per day,fig.width=5}
hist(StepsPerDay,ylim=c(0,40),main="Average steps per day",xlab = "Steps/day")
```

### Mean & Median of total steps per day

```{r mean-median totalSteps per day}
mean(StepsPerDay,na.rm = TRUE)
median(StepsPerDay,na.rm = TRUE)
```

<div style="margin-bottom:30px;">
</div>


## **Steps per interval**

### Plot of steps per 5-min interval averaged accross all days

```{r steps per interval plot}
stepsperinterval<-ActivityData%>%
                    group_by(interval)%>%
                      summarise(steps=mean(steps,na.rm = TRUE))
plot(stepsperinterval$interval,stepsperinterval$steps,type="l",col="red",lwd=1.5,xaxt="n",xlab = "Interval Time",ylab="Steps",main = "Average steps per 5-min interval accross all days")
axis(side=1, at=seq(0,max(stepsperinterval$interval), by=100))

```

### Interval with max no. of steps

```{r interval with max no. of steps}
stepsperinterval[which.max(stepsperinterval$steps),1][[1]]
```

<div style="margin-bottom:30px;">
</div>


## **Handling missing values**

### Total missing observations

```{r total missing observations}
sum(is.na(ActivityData))
```


### Replacing missing steps values, with its intervals mean

```{r replacing missing values}
for (i in 1:length(ActivityData$steps)){
      if (is.na(ActivityData$steps[i]))
        ActivityData$steps[i]= stepsperinterval[match(ActivityData$interval[i],stepsperinterval$interval),2]
ActivityData$steps<-unlist(ActivityData$steps)
}
  
```

### New steps per day histogram

```{r hist totalSteps per day with imputed data,fig.width=5}
StepsPerDaynew<-tapply(ActivityData$steps,ActivityData$date,sum)
hist(StepsPerDaynew,ylim=c(0,40),main="Average steps per day with imputed values",xlab="Steps/day")
```

### New mean of steps per day

```{r mean totalSteps per day with imputed data}
mean(StepsPerDaynew,na.rm = TRUE)
```

### New median of steps per day
```{r median totalSteps per day with imputed data}
median(StepsPerDaynew,na.rm = TRUE)
```
The new median is `r format(median(StepsPerDaynew,na.rm = TRUE), scientific=F)` , a very small change from previous median which was `r format(median(StepsPerDay,na.rm = TRUE), scientific=F)`. The mean is the same.The impact of imputing missing data on the estimates of the total daily number of steps, is minimal in this case.


<div style="margin-bottom:30px;">
</div>


## **Weekdays vs Weekends**
### Creating a new factor variable "Day" in the dataset

```{r Day}

ActivityData$date<-as.Date(ActivityData$date)
weekday<-weekdays(ActivityData$date)

for(i in 1:length(weekday)){
  if(weekday[i]=="Saturday"| weekday[i]=="Sunday"){
    weekday[i]<-"weekend"
  } else {
    weekday[i]<-"weekday"
   }
}   
  
ActivityData$Day<-weekday
ActivityData$Day<-as.factor(ActivityData$Day)
head(ActivityData)
```

### Plot depicting average steps for Weekend and Weekdays, per 5-minute intervals
```{r plot Weekdays vs Weekends,fig.height=7,fig.width=8}
ActivityData$interval<-as.numeric(ActivityData$interval)

StepsperWeekDay<-ActivityData%>%
  group_by(interval,Day)%>%
  summarise(steps=mean(steps))

library(ggplot2)

ggplot(StepsperWeekDay,aes(interval,steps))+
  facet_wrap(~Day,nrow=2,ncol=1)+
  scale_x_continuous(breaks=seq(0,max(StepsperWeekDay$interval),200))+
  geom_line(color="red")+
  ylab("Steps Averaged across days")+
  xlab("Interval time")+
  labs(title = "Steps per Weekday & Weekend, by interval time",size=2,
       subtitle = "Weekdays steps exceed Weekends steps")+
  theme( strip.text=element_text(size=12,color="dodgerblue4"))+
theme(plot.title=element_text(size=18, hjust=0.5, face="bold", colour="black"))+
theme(plot.subtitle=element_text(size=14, hjust=0.5, face="italic", color="chocolate"))

```
